---
title: "tidy data exercises (factors and dates)"
format: html
editor: source
---



```{r}
library(tidyverse)
```

# Data formats


## Example 1:
Have a look at the following dataset:



```{r}
rating_data <- data.frame(
  choice = c("Strongly disagree", "disagree", "neither agree nor disagree", "agree", "strongly agree"),
  freq = c(10, 12, 13, 50, 21)
)

rating_data
```


Create a bar plot showing the frequency of each choice. 

```{r}
ggplot(rating_data, aes(choice, freq)) + geom_bar(stat = 'identity') + coord_flip()
```

What's the issue?

How can you fix this?


## Example 2:


```{r}
load("datasets/date_example.RData")

date_example


```


### Exercise
Create a figure showing rate as function of date1; use both points and lines.

```{r}

```


What can you say about this figure?




### Exercise
Create a figure showing rate as function of date2; use both points and lines.

```{r}

```


What can you say about this figure?




### Exercise
Create a figure showing rate as function of date3; use both points and lines.

```{r}

```


What can you say about this figure?


## How to encode dates?

Have a look at the [lubridate sheatsheet](https://rstudio.github.io/cheatsheets/lubridate.pdf)




