---
title: "tidy data exercises (renaming)"
format: html
editor: source
---



```{r}
library(tidyverse)
```

# Tidyng data

```{r}

# example from:
# https://rstudio-pubs-static.s3.amazonaws.com/931128_d1b99faddf5c45e098a797c8d6b9c5a3.html
dset <- as.data.frame(matrix(rnorm(n = 70, 0, 1), ncol = 7))
names(dset) <- c("first column", "SEC0nd!", " % valueColumn","SAME NAME","SAME NAME", "Age", "$Salary$") 
dset
```



## Exercises


```{r}
# compute the mean of the vrialble called "first column"


```



```{r}
# compute the median of the vrialble called "$Salary$"


```



```{r}
# create a new variable that sums "SEC0nd!" and "$Salary$"


```




What do you thinnk about those column names?

How can you change them?





## How to chose good variable names (column names)

For a long version see:\


https://behaverse.org/data-model/explanations/naming.html


For a short version:

-   use snake_case ("math_grades" MathGrade Mathgrades "Math Grades")
-   use variable names that are meaningful (i.e., stand on their own) "population" "population_count" "pop" "qst_1"
-   don't use abreviations (population_count, pop_n, )
-   you must be consistent;


## How to rename columns?

### dplyr

```{r}

```



### janitor

```{r}
library(janitor)
```




