---
title: "Basic statistics using tidyverse"
format: html
---



This notebook provides exercises for simple and multiple linear regression.
The data used here comes from the ISLR package that you will need to install first.
We will also use the broom package.





```{r}
# load packages
library(tidyverse)
library(ISLR)
library(report)
```




```{r}
data(Wage)
head(Wage)
```


## Statistics


### Correlation

Comptue the correlation between wage and age.

```{r}


```

What can you say about that correlation?



Can you get access to the p-value and store it in a variable? 
 





### T-test

Do people with Very Good health have higher wages than those with worse health?

```{r}

```


What can you conclude?

Can you get access to the p_value and store it as a variable?


## The "report" package

Sometimes it might be tricky to accurately report the results of an analysis. The `report` package can help!

```{r}
cor_result <- cor.test(Wage$wage, Wage$age)
report(cor_result)

```


It works with many statistical tests. Try it on the t-test that was computed earlier.


```{r}

```


## The "broom" package

```{r}
library(broom)
```


```{r}

t_test_result <- t.test(Wage$wage~Wage$health)
broom::tidy(t_test_result)

```



## Compute stats the tidyverse way

`summarise()` requires each expression to return a single value, so it cannot
handle `broom::tidy()` results (which are tibbles with multiple columns). Use
`reframe()` instead — it is designed for exactly this case and works seamlessly
with `group_by()`.

```{r}
# t.test via reframe — works on the whole dataset
Wage |>
  reframe(broom::tidy(t.test(wage ~ health)))
```

```{r}
# cor.test via reframe
Wage |>
  reframe(broom::tidy(cor.test(wage, year)))
```




What's great about this approach is that we can now use the power of tidyverse. 

For example, image we want to compute a t-test of wage for two levels of health (as above) but separately for each level of education. 

In base R, this would involve quite a lot of coding; with the tidyverse and broom, this is becomes a breeze.


In the chunck below, update the code we wrote above to compute the t-tests and correlations tests separetly for each level of education.



```{r}

```



